Network Configuration Synthesis with Abstract Topologies

### 2016-11-12 paper/abstraction/paper.tex
minor edits

\newtheorem{thm}{Theorem}
\newtheorem{defn}{Definition}
\newtheorem{lem}[thm]{Lemma}

@@

\subsection{Compilation}
\label{subsec:compilation}

@@

%Although \sysname can ensure correct forwarding behavior independent of the exact concrete topology or failure scenario, it is often useful to be able to ask "what if" questions about other network properties. For example, an operator may require a certain level fault tolerance between between different collections of routers. Such analysis depends, not only on the network topology, but also on the underlying routing policy, both of which \sysname has access to.
The \sysname compiler generates templates from the inputs above in three phases. First, it combines the abstract topology and routing policy into a {\em product graph} (PG) that compactly captures the flow of routing information over the topology in a policy-compliant manner. The algorithm for this phase builds on the Propane compiler, which was designed for concrete topologies and policy; we show that it can be extended to abstract inputs and that the extensions are correct.

@@

The PG compactly encodes all paths protocol messages should be passed along through the network in order to guarantee the devices find policy-compliant paths dynamically.
%
%Although \sysname policies describe the routes that traffic should take through the network, network protocols (including BGP) typically disseminate router information in the opposite direction. For example, a router Y will tell its neighbor X that it has a route to the destination meaning that X can send traffic through Y. For this reason, we start by compiling
%
Because \sysname policies can have a separate routing policy for each predicate matching traffic, we construct one such PG for each predicate.

@@

The PG is constructed by adding an edge from state $m = (l_m, q_{m_1}, \dots, q_{m_k})$ to $n = (l_n, q_{n_1}, \dots, q_{n_k})$ whenever $\sigma_i(q_{m_i}, l_n) = q_{n_i}$ for each $i$ and $(l_m,l_n) \in E$ is a valid topology link.
%
We add edges from the start node to any $m = (l, q_{m_1}, \dots, q_{m_k})$ when $\sigma_i(q_{0_i}, l) = q_{m_i}$ for each $i$.
%
The preference function $P(m)$ denotes the rank of paths through the PG ending at node $m$ and is defined as $P(m) = \set{i~\vert~q_{m_i} \in F_i}$.
%
Finally, we write $\tilde{m} = l$ to extract the topology location from a PG node, when $m = (l, q_{m_1}, \dots, q_{m_k}) \in V'$.

Figure~\ref{fig:example-compilation} shows an example of a PG for the policy that applies to all external traffic from the datacenter example in \S\ref{sec:overview} ($\True ~\Path~ \CD{\Exit(N1 \Prefer~N2)}$).
%
%\begin{code}
%\True \Path \Exit(\sf{Peer1}) \Prefer \Exit(\sf{Peer2})
%\end{code}
%\noindent%
%
The first automaton represents the more preferred constraint $\CD{\Exit(N1)}$ and the second automaton represents the less preferred constraint $\CD{\Exit(N2)}$. The PG is shown for both an instance of a simple concrete network matching the abstraction from \S\ref{sec:overview} as well as for the abstract topology.

@@

\begin{defn}
If we have a graph homomorphism $f : G \rightarrow G^A$, concrete $PG = (G',\mathit{start},P)$ and abstract $PG^A = (G'^A, \mathit{start}^A, P^A)$, then we can construct a new homomorphism $f_{pg} : G' \rightarrow G'^A$ as follows:
\[ \begin{array}{rcl}
  f_{pg}( \mathit{start} ) & = & \mathit{start}^A  \\
  f_{pg}( (l,q_1,\ldots,q_n) ) & = & (f(l),q_1,\ldots,q_n) \\
\end{array} \]
\end{defn}

@@

\section{Template Generation}

@@

Lemma~\ref{lem:preference} gives us reason to believe that we can perform compilation over the abstract PG directly and ensure that the inferred device-level preferences will be correct when instantiating the template with a concrete network.

@@

\begin{thm}
  For any context $\Gamma$, topology $G$, abstract topology $G^A$, homomorphism $f : G \rightarrow G^A$, and policy $pol$,

  \noindent
  $\Con(\text{compile}(pol,G^A), \Gamma,f,G) = \text{compile}(\Con(pol,\Gamma,f),G)$
  \label{thm:concretization}
\end{thm}
%
\vspace{-1em}
Full definitions of compilation and concretization, as well as the proof of Theorem~\ref{thm:concretization} are included in the appendix.

@@

To evaluate \sysname, we apply it on policies for backbone and datacenter networks. We aim to evaluate the expressiveness of our abstracttions, the precision of the abstract disjoint-path analysis, and the compilation time of \sysname both with abstraction and without.

@@

\subsection{Compilation time}

\begin{figure}[t!]
  \subcaptionbox{Datacenter}
    {\includegraphics[width=.49\columnwidth]{figures/Fattree-time.png}}
  \subcaptionbox{Backbone}
    {\includegraphics[width=.49\columnwidth]{figures/backbone-time.png}} \\
  \caption{Concrete vs. Abstract Compilation Time. \label{fig:compilation-times}}
  \vspace{-1em}
\end{figure}

\begin{figure}[t!]
  \subcaptionbox{Datacenter}
    {\includegraphics[width=.49\columnwidth]{figures/Fattree-analysis-time.png}}
  \subcaptionbox{Backbone}
    {\includegraphics[width=.49\columnwidth]{figures/backbone-analysis-time.png}} \\
  \caption{Abstract Compilation Time by Phase. \label{fig:abstract-breakdown}}
  \vspace{-1em}
\end{figure}

Using abstraction allows us to naturally scale network synthesis in \sysname. We evaluate compilation time in \sysname both with and without abstraction using routing policy for backbone and datacenter networks inspired by configurations obtained from a large cloud provider. For both types networks, we fix routing policy but scale the size of the underlying topology.

@@

Figure~\ref{fig:compilation-times} shows compilation time for the same policies for both the abstract and concrete versions of the policies described above. All experiments are run on an 8 core, 2.4 GHz Intel i7 processor Mac with 8GB of Ram, and the time taken is the average of 3 runs.

For both the networks, the abstract compilation starts off slightly slower than concrete compilation for small topologies due to the overhead of the fault-tolerance analysis. However, as the size of the topology increases, abstract compilation becomes orders of magnitude faster than concrete compilation. In all cases for both networks, abstract compilation takes less than 10 seconds to complete.

Figure~\ref{fig:abstract-breakdown} shows the relative time taken by each phase of abstract compilation. Notably, the fault-tolerance analysis takes the majority of the time, however the analysis does not depend on the number of concrete nodes in the network, and thus is largely a fixed cost. In particular, the number of calls to $\nu$Z remains constant the across topology size. The seesaw behavior for the data center networks results from differences in time taken by $\nu$Z to minimize similar constraints with different values.

>>>>

\newtheorem{thm}{Theorem}[section]
\newtheorem{defn}{Definition}[section]
\newtheorem{lem}[thm]{Lemma}

@@

\subsection{Template Generation}
\label{subsec:compilation}

@@

%Although \sysname can ensure correct forwarding behavior independent of the exact concrete topology or failure scenario, it is often useful to be able to ask "what if" questions about other network properties. For example, an operator may require a certain level fault tolerance between between different collections of routers. Such analysis depends, not only on the network topology, but also on the underlying routing policy, both of which \sysname has access to.
The \sysname compiler generates templates from the inputs above in three phases. First, it combines the abstract topology and routing policy into a {\em product graph} (PG) that compactly captures the flow of routing information over the topology in a policy-compliant manner. The algorithm for this phase builds on the \propane compiler, which was designed for concrete networks; we show that it can be extended to abstract inputs and that the extensions are correct.

@@

The PG compactly encodes all paths that protocol messages should be passed along through the network in order to guarantee that the devices find policy-compliant paths dynamically.
%
%Although \sysname policies describe the routes that traffic should take through the network, network protocols (including BGP) typically disseminate router information in the opposite direction. For example, a router Y will tell its neighbor X that it has a route to the destination meaning that X can send traffic through Y. For this reason, we start by compiling
%
Because \sysname policies can have a separate routing policy for each predicate matching traffic, we construct one such PG for each predicate.

@@

The PG is constructed by adding an edge from state $m = (l_m, q_{m_1}, \dots, q_{m_k})$ to $n = (l_n, q_{n_1}, \dots, q_{n_k})$ whenever $\sigma_i(q_{m_i}, l_n) = q_{n_i}$ for each $i$ and $(l_m,l_n) \in E$.
%
We add edges from the $\mathit{start}$ node to any $m = (l, q_{m_1}, \dots, q_{m_k})$ when $\sigma_i(q_{0_i}, l) = q_{m_i}$ for each $i$.
%
The preference function $P(m)$ denotes the rank of paths through the PG ending at node $m$ and is defined as $P(m) = \set{i~\vert~q_{m_i} \in F_i}$.
%
Finally, we write $\tilde{m} = l$ to extract the topology location from a PG node, when $m = (l, q_{m_1}, \dots, q_{m_k}) \in V'$.

Figure~\ref{fig:example-compilation} shows an example of a PG for the policy that applies to all external traffic from the datacenter example in \S\ref{sec:overview}.
%
%\begin{code}
%\True \Path \Exit(\sf{Peer1}) \Prefer \Exit(\sf{Peer2})
%\end{code}
%\noindent%
%
The first automaton represents the more preferred constraint $\CD{\Exit(N1)}$ and the second automaton represents the less preferred constraint $\CD{\Exit(N2)}$. The PG is shown for both an instance of a simple concrete network matching the abstraction from \S\ref{sec:overview} as well as for the abstract topology.

@@

\begin{defn}
\label{def:homomorphism}
If we have a graph homomorphism $f : G \rightarrow G^A$, concrete $PG = (G',\mathit{start},P)$ and abstract $PG^A = (G'^A, \mathit{start}^A, P^A)$, then we can construct a new homomorphism $f_{pg} : G' \rightarrow G'^A$ as follows:
\[ \begin{array}{rcl}
  f_{pg}( \mathit{start} ) & = & \mathit{start}^A  \\
  f_{pg}( (l,q_1,\ldots,q_n) ) & = & (f(l),q_1,\ldots,q_n) \\
\end{array} \]
\end{defn}

As we will see in \S\ref{sec:generation}, we have proven that our generation strategy commutes with concrete expansion, meaning that we obtain the same results if we expand the abstraction early, or if we defer the expansion until after template generation.
%
Definition~\ref{def:homomorphism} provides some intuition concerning why our strategy for generating templates is correct.
%Abstract product graphs are related to concrete product graphs via a graph homomorphism, just as abstract topologies are related to concrete topologies via a graph homomorphism.

@@

\section{Template Generation}
\label{sec:generation}

@@

Lemma~\ref{lem:preference} tells us that inferring preferences over the abstract PG before concrete expansion is equivalent to inferring preferences over an already-expanded concrete PG.

@@

\begin{thm}
  For any context $\Gamma$, topologies $G$ and $G^A$, homomorphism $f : G \rightarrow G^A$, and policy $pol$,
  $$\Con(\text{compile}(pol,G^A), \Gamma,f,G) = \text{compile}(\Con(pol,\Gamma,f),G)$$
  \label{thm:concretization}
\end{thm}
%
\vspace{-1em}
Full definitions of compilation and concretization, as well as the proof of Theorem~\ref{thm:concretization} are included in the appendix.

@@

To evaluate \sysname, we apply it on policies for backbone and datacenter networks. We aim to evaluate the expressiveness of our abstracttions, the precision of the abstract disjoint-path analysis, and the generation time of \sysname both with abstraction and without.

@@

\subsection{Generation time}

\begin{figure}[t!]
  \subcaptionbox{Datacenter}
    {\includegraphics[width=.49\columnwidth]{figures/Fattree-time.png}}
  \subcaptionbox{Backbone}
    {\includegraphics[width=.49\columnwidth]{figures/backbone-time.png}} \\
  \caption{Concrete vs. Abstract Generation Time. \label{fig:compilation-times}}
  \vspace{-1em}
\end{figure}

\begin{figure}[t!]
  \subcaptionbox{Datacenter}
    {\includegraphics[width=.49\columnwidth]{figures/Fattree-analysis-time.png}}
  \subcaptionbox{Backbone}
    {\includegraphics[width=.49\columnwidth]{figures/backbone-analysis-time.png}} \\
  \caption{Abstract Generation Time by Phase. \label{fig:abstract-breakdown}}
  \vspace{-1em}
\end{figure}

Using abstraction allows us to naturally scale network synthesis in \sysname. We evaluate generation time in \sysname both with and without abstraction using routing policy for backbone and datacenter networks inspired by configurations obtained from a large cloud provider. For both types networks, we fix routing policy but scale the size of the underlying topology.

@@

Figure~\ref{fig:compilation-times} shows generation time for the same policies for both the abstract and concrete versions of the policies described above. All experiments are run on an 8 core, 2.4 GHz Intel i7 processor Mac with 8GB of Ram, and the time taken is the average of 3 runs.

For both the networks, the abstract generation starts off slightly slower than concrete generation for small topologies due to the overhead of the fault-tolerance analysis. However, as the size of the topology increases, abstract generation becomes orders of magnitude faster than concrete generation. In all cases for both networks, abstract generation takes less than 10 seconds to complete.

Figure~\ref{fig:abstract-breakdown} shows the relative time taken by each phase of abstract generation. Notably, the fault-tolerance analysis takes the majority of the time, however the analysis does not depend on the number of concrete nodes in the network, and thus is largely a fixed cost. In particular, the number of calls to $\nu$Z remains constant the across topology size. The seesaw behavior for the data center networks results from differences in time taken by $\nu$Z to minimize similar constraints with different values.
